---
title: "PCLO and Hepatocellular Carcinoma Code Supplement"
author: "Ariana Chriss, Tri Dong, Paul Morgan, Alec Pankow, Bryce Rowan, Rozalyn Wood"
date: "February 20, 2022"
output:
  html_document:
    df_print: paged
subtitle: BSR1026 Group 1
---

Organized by primary and secondary analyses carried out by each team member. 

### Bryce

```{r}
#This script gets through the analyses I used and includes the imputation set up for the data used in the midterm project
#Step 1: Read in files and impute PCLO mutation
library(openxlsx)
Midterm_Dataset <- read.xlsx("Midterm_Dataset.xlsx")
Keep_SubsetA <- Midterm_Dataset[,c(1:3,7:13,19,22)]
Keep_SubsetB <- as.data.frame(Keep_SubsetA)
Mutated <- Keep_SubsetB[which(Keep_SubsetB$PCLO == "Mutated"),]
Not_Mutated <- Keep_SubsetB[which(Keep_SubsetB$PCLO == "not mutated"),]
Extract_Missing_PCLO <- Keep_SubsetB[!complete.cases(Keep_SubsetB$PCLO),]
nrow(Extract_Missing_PCLO)
Extract_Missing_PCLO$PCLO <- "not mutated"
Extract_Missing_PCLO$PCLO
Pull_Full_PCLO <- Keep_SubsetB[complete.cases(Keep_SubsetB$PCLO),]
nrow(Pull_Full_PCLO)
New_Full_PCLO <- rbind(Pull_Full_PCLO,Extract_Missing_PCLO)
table(New_Full_PCLO$PCLO)
addmargins(table(New_Full_PCLO$PCLO,dnn=('PCLO')))
summary(New_Full_PCLO)
Mutated <- New_Full_PCLO[which(New_Full_PCLO$PCLO == "Mutated"),]#Updated version
Not_Mutated <- New_Full_PCLO[which(New_Full_PCLO$PCLO == "not mutated"),]#Updated version
#Now to impute the rest of the dataset based on our agreed upon rules
Healthy_Full <- New_Full_PCLO[which(New_Full_PCLO$BMI_Category == "Healthy Weight"),]
Overweight_Full <- New_Full_PCLO[which(New_Full_PCLO$BMI_Category == "Overweight"),]
Underweight_Full <- New_Full_PCLO[which(New_Full_PCLO$BMI_Category == "Underweight"),]
Obesity_Full <- New_Full_PCLO[which(New_Full_PCLO$BMI_Category == "Obesity"),]
Missing_Weight_Full <- New_Full_PCLO[!complete.cases(New_Full_PCLO$BMI_Category),]
Missing_Weight_Full$BMI_Category <- "Healthy Weight"
Upgraded_Healthy_Full <- rbind(Healthy_Full,Missing_Weight_Full)
nrow(Upgraded_Healthy_Full) - nrow(Healthy_Full)
BMI_imputed_New_Full_PCLO <- rbind(Obesity_Full,Overweight_Full,Underweight_Full,Upgraded_Healthy_Full)
nrow(BMI_imputed_New_Full_PCLO)
summary(as.factor(BMI_imputed_New_Full_PCLO$Alcohol_History))
Missing_AlcHis <- BMI_imputed_New_Full_PCLO[!complete.cases(BMI_imputed_New_Full_PCLO$Alcohol_History),]
AlcHis <- BMI_imputed_New_Full_PCLO[complete.cases(BMI_imputed_New_Full_PCLO$Alcohol_History),]
Missing_AlcHis$Alcohol_History <- 0
HisAlc_BMI_imputed_New_Full_PCLO <- rbind(AlcHis, Missing_AlcHis)
summary(as.factor(HisAlc_BMI_imputed_New_Full_PCLO$Sex))
#No missing values for sex
summary(as.factor(HisAlc_BMI_imputed_New_Full_PCLO$Hepatitis_History))
Missing_HepHis <- HisAlc_BMI_imputed_New_Full_PCLO[!complete.cases(HisAlc_BMI_imputed_New_Full_PCLO$Hepatitis_History),]
HepHis <- HisAlc_BMI_imputed_New_Full_PCLO[complete.cases(HisAlc_BMI_imputed_New_Full_PCLO$Hepatitis_History),]
Missing_HepHis$Hepatitis_History <- 0
HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO <- rbind(HepHis,Missing_HepHis)
Missing_Mutation_Burden <- HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO[!complete.cases(HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO$Mutation_Burden),]
Mutation_Burden <- HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO[complete.cases(HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO$Mutation_Burden),]
Missing_Mutation_Burden$Mutation_Burden <- 112.4
BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO <- rbind(Mutation_Burden,Missing_Mutation_Burden)
#Save the files for group (Assumes you have a folder created to store the files)
#save(BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO, file = "Additional_Files_rdata/Main_variables_imputed_dataset_MidtermProject_02_13_22.RData")
Liver_Fibrosis <- BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO[complete.cases(BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO$Liver_Fibrosis),]
#save(Liver_Fibrosis, file = "Additional_Files_rdata/LiverFibrosis_Subset_Complete_Dataset_MidtermProject_02_13_22")
Recurrence <- BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO[complete.cases(BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO$Recurrence),]
#save(Recurrence, file = "Additional_Files_rdata/Recurrence_Subset_Complete_Dataset_MidtermProject_02_13_22") 
#Use pairwise wilcox test for the secondary level analyses
pairwise.wilcox.test(BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO$Mutation_Burden,BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO$BMI_Category,p.adjust.method="fdr")
pairwise.wilcox.test(BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO$Mutation_Burden,BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO$BMI_Category,p.adjust.method="none")
#Output from test is a matrix of pvalues; fdr is a form of correcting for multiple comparisons in testing with other methods
#Trends still hold, but the method did what was suppossed to do!
#Have to use continous outcome with categorical variable
Females <- BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO[which(BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO$Sex == "FEMALE"),]
Males <- BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO[which(BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO$Sex == "MALE"),]
summary(Females)
summary(Males)
var.test(Females$Age, Males$Age, alternative = "two.sided")
t.test(Females$Age, Males$Age, var.equal=TRUE)
table(Females$PCLO,Females$Age)
chisq.test(table(Females$PCLO,Females$Age))
fisher.test(table(Females$PCLO,Females$Age))
table(Males$PCLO,Males$Age)
#Grouping attempt
Mt_Females <- Females[which(Females$PCLO == "Mutated"),]
Norm_Females <- Females[which(Females$PCLO == "not mutated"),]
Norm_Males <- Males[which(Males$PCLO == "not mutated"),]
Mt_Males <- Males[which(Males$PCLO == "Mutated"),]
CancerbySexG <- rbind(Mt_Females,Mt_Males,Norm_Females,Norm_Males)
Mt_Males$NewGroup <- "Mutated_Males"
Mt_Females$NewGroup <- "Mutated_Females"
Norm_Females$NewGroup <- "Norm_Females"
Norm_Males$NewGroup <- "Norm_Males"
CancerbySexG <- rbind(Mt_Females,Mt_Males,Norm_Females,Norm_Males)
pairwise.wilcox.test(CancerbySexG$Age, CancerbySexG$NewGroup, p.adjust.method = "fdr")
pairwise.wilcox.test(CancerbySexG$Age, CancerbySexG$NewGroup, p.adjust.method = "none")
#Now to find the median and the IQR for the signficant results from the pairwisewilcoxon rank sum test
median(Norm_Females$Age)
IQR(Norm_Females$Age)
median(Norm_Males$Age)
IQR(Norm_Males$Age)
#Adding in information about the t-test between PCLO mutation or lack thereof and age
nrow(Mutated)
nrow(Not_Mutated)
var.test(Mutated$Age, Not_Mutated$Age, alternative= "two.sided")
t.test(Mutated$Age, Not_Mutated$Age, var.equal=TRUE)
#Adding information on boxplot generation for figures
boxplot(Age~NewGroup, data=CancerbySexG, col=(c("mediumorchid3","slateblue3")), main = "PCLO by Sex Over Age", xlab="PCLO by Sex", ylab = "Age")
boxplot(Mutation_Burden~NewGroup, data=CancerbySexG, col=(c("mediumorchid3","slateblue3")), main = "PCLO by Sex Over Mutation Burden", xlab="PCLO by Sex", ylab = "Mutation Burden")
#Above were the original plots. Now we are adding in the plots with the resolution everyone wants
png(file="testdrive5_PCLO_by_Sex_Over_Age.png", width = 1000, height = 1000); par(mar = c(5,5,2,2), xaxs = "i", yaxs = "i", cex.axis=2, cex.lab= 2); boxplot(Age~NewGroup, data=CancerbySexG, col=(c("mediumorchid3","slateblue3")), main = "PCLO by Sex Over Age", xlab="PCLO by Sex", ylab = "Age"); legend("bottomleft", legend = c("Female", "Male"), fill = c("mediumorchid3","slateblue3"), pch = 5, cex=2); dev.off()
png(file="testdriveB_PCLO_by_Mutation_Burden_Over_Age.png", width = 1000, height = 1000); par(mar = c(5,5,2,2), xaxs = "i", yaxs = "i", cex.axis=2, cex.lab= 2); boxplot(Mutation_Burden~NewGroup, data=CancerbySexG, col=(c("mediumorchid3","slateblue3")), main = "PCLO by Sex Over Mutation Burden", xlab="PCLO by Sex", ylab = "Mutation Burden"); legend("topleft", legend = c("Female", "Male"), fill = c("mediumorchid3","slateblue3"), pch = 5, cex=2); dev.off()
#The colors are seperated by gender
#Additional tests to try
chisq.test(table(Males$PCLO,Males$BMI_Category))
fisher.test(table(Males$PCLO,Males$BMI_Category))
chisq.test(table(Males$PCLO,Males$Recurrence))
chisq.test(table(Males$PCLO,Males$Metastasis))
chisq.test(table(Males$PCLO,Males$Hepatitis_History))
chisq.test(table(Males$PCLO,Males$Liver_Fibrosis))
chisq.test(table(Males$PCLO,Males$Survival_Status))
chisq.test(table(Males$PCLO,Males$Alcohol_History))
chisq.test(table(Females$PCLO,Females$Alcohol_History))
fisher.test(table(Females$PCLO,Females$Alcohol_History))
chisq.test(table(Females$PCLO,Females$Metastasis))
fisher.test(table(Females$PCLO,Females$Metastasis))
fisher.test(table(Females$PCLO,Females$Recurrence))
fisher.test(table(Females$PCLO,Females$Survival_Status))
fisher.test(table(Females$PCLO,Females$Hepatitis_History))
fisher.test(table(Females$PCLO,Females$Liver_Fibrosis))
pairwise.wilcox.test(CancerbySexG$Mutation_Burden, CancerbySexG$NewGroup, p.adjust.method = "none")
pairwise.wilcox.test(CancerbySexG$Mutation_Burden, CancerbySexG$NewGroup, p.adjust.method = "fdr")
#Method to create table one, requires installing package, followed by loading it
library(tableone)
ncol(BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO)
TempFile <- BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO[,-1]
ncol(TempFile)
colnames(TempFile)
#Subset worked!
#Going to try to make the other categorical variables categorical
Second_TempFile <- TempFile
Second_TempFile$Hepatitis_History <- as.factor(Second_TempFile$Hepatitis_History)
Second_TempFile$Survival_Status <- as.factor(Second_TempFile$Survival_Status)
Second_TempFile$Recurrence <- as.factor(Second_TempFile$Recurrence)
Second_TempFile$Metastasis <- as.factor(Second_TempFile$Metastasis)
Second_TempFile$Liver_Fibrosis <- as.factor(Second_TempFile$Liver_Fibrosis)
CreateTableOne(data=Second_TempFile)
#Now to do the same for the CancerbySexG dataset I made
Second_Temp_CancerbySexG <- CancerbySexG[,-1]
Second_Temp_CancerbySexG$Hepatitis_History <- as.factor(Second_Temp_CancerbySexG$Hepatitis_History)
Second_Temp_CancerbySexG$Survival_Status <- as.factor(Second_Temp_CancerbySexG$Survival_Status)
Second_Temp_CancerbySexG$Recurrence <- as.factor(Second_Temp_CancerbySexG$Recurrence)
Second_Temp_CancerbySexG$Metastasis <- as.factor(Second_Temp_CancerbySexG$Metastasis)
Second_Temp_CancerbySexG$Liver_Fibrosis <- as.factor(Second_Temp_CancerbySexG$Liver_Fibrosis)
CreateTableOne(data=Second_Temp_CancerbySexG)
```

### Rozalyn

```{r, eval=F, warning=F}
library(readxl)
PcloDS<-read_xlsx("data/Midterm_Dataset.xlsx")
library(dplyr)

#Primary Outcomes

#Mutation Burden + Gene (mutation/non mutation)
#First: Subset data to Mutation Burden+ PCLO with and without mutation
PcloMtN<-subset(x=PcloDS,
             subset=PCLO=="not mutated",
             select=c("Mutation_Burden","PCLO"))

PcloMtY<-subset(x=PcloDS,
                subset=PCLO=="Mutated",
                select=c("Mutation_Burden","PCLO"))

#Test statistics for Mutation burden + PCLO using a two sample t-test
PcloMt1<-subset(x=PcloDS,
               subset=PCLO=="Mutated",
               select=c("Mutation_Burden"))

PcloMt0<-subset(x=PcloDS,
                subset=PCLO=="not mutated",
                select=c("Mutation_Burden"))

#Impute missing values with Mean for mutation burden
PcloMt0$Mutation_Burden[is.na(PcloMt0$Mutation_Burden)] <- mean(PcloMt0$Mutation_Burden, na.rm = TRUE)
PcloMt1$Mutation_Burden[is.na(PcloMt1$Mutation_Burden)] <- mean(PcloMt1$Mutation_Burden, na.rm = TRUE)

# F-Test for Variance
require(stats)
var.test(x=PcloMt1$Mutation_Burden, y=PcloMt0$Mutation_Burden, alternative="two.sided", na.action=TRUE)

#Using Welch's Two sample t-test because variance is unequal (based on F-test above)
t.test(x=PcloMt1,
       y=PcloMt0,
       alternative ="two.sided",
       paired=FALSE, 
       var.equal=FALSE,
       conf.level=0.95)

#Age by PCLO mutation status 
#Subset data first
PcloAge1<-subset(x=PcloDS,
                subset=PCLO=="Mutated",
                select=c("Age"))

PcloAge0<-subset(x=PcloDS,
                subset=PCLO=="not mutated",
                select=c("Age"))

#Impute missing values with Mean for mutation burden
PcloAge0$Age[is.na(PcloAge0$Age)] <- mean(PcloAge0$Age, na.rm = TRUE)
PcloAge1$Age[is.na(PcloAge1$Age)] <- mean(PcloAge1$Age, na.rm = TRUE)

#Secondary Outcomes

#Liver fibrosis + PCLO - Basic association between gene mutation and LF
#First Subset Data
PcloLFN<-subset(x=PcloDS,
                subset=PCLO=="Mutated",
                select=c("Liver_Fibrosis","PCLO"))

PcloLFY<-subset(x=PcloDS,
                subset=PCLO=="not mutated",
                select=c("Liver_Fibrosis","PCLO"))

PcloLFNNA<-na.omit(PcloLFN)
PcloLFYNA<-na.omit(PcloLFY)

PcloLFAll<-rbind(PcloLFNNA,PcloLFYNA)

PcloLFAll$PCLO[PcloLFAll$PCLO=="Mutated"]<-"1"
PcloLFAll$PCLO[PcloLFAll$PCLO=="not mutated"]<-"0"

#Test statistic for PCLO +LF

#checking counts
length(PcloLFY[PcloLFY=='0'])
length(PcloLFY[PcloLFY=='1'])
length(PcloLFN[PcloLFN=='0'])
length(PcloLFN[PcloLFN=='1'])

#Creating 2 by 2 dataframe from counts above
PcloLFTable<-data.frame(
  "Mutated PCLO" = c(8,14),
  "Not Mutated PCLO" = c(68,122),
  row.names = c("No Liver Fibrosis", "Liver Fibrosis"),
  stringsAsFactors = FALSE
)

#Fisher exact test
fisher.test(PcloLFTable,
            conf.level = 0.95,
            alternative="two.sided")
#Outcome=Not significant based on fisher exact with approximately 50% of LF data NA

#Secondary Outcomes with 3 Variables
#Liver Fibrosis + Mutation Burden + PCLO Mutation Status: Subset data into 4 groups divided by PCLO mutation status + Liver Fibrosis
load("LiverFibrosisDS.RData")
LF_MB_P0<-subset(x=LiverFibrosis_completeOnly_rest_FullSet,
              subset=PCLO=="not mutated",
              select=c("Mutation_Burden","Liver_Fibrosis"))

LF_MB_P1<-subset(x=LiverFibrosis_completeOnly_rest_FullSet,
                 subset=PCLO=="Mutated",
                 select=c("Mutation_Burden","Liver_Fibrosis"))
LF0_MB_P0<-subset(x=LF_MB_P0,
                 subset=Liver_Fibrosis=="0",
                 select=c("Mutation_Burden"))

LF1_MB_P0<-subset(x=LF_MB_P0,
                  subset=Liver_Fibrosis=="1",
                  select=c("Mutation_Burden"))

LF0_MB_P1<-subset(x=LF_MB_P1,
                  subset=Liver_Fibrosis=="0",
                  select=c("Mutation_Burden"))

LF1_MB_P1<-subset(x=LF_MB_P1,
                  subset=Liver_Fibrosis=="1",
                  select=c("Mutation_Burden"))

#Creating column for groups 
LF0_MB_P0$LF_PCLO<-"PCLO_0/LF_0"
LF1_MB_P0$LF_PCLO<-"PCLO_0/LF_1"
LF0_MB_P1$LF_PCLO<-"PCLO_1/LF_0"
LF1_MB_P1$LF_PCLO<-"PCLO_1/LF_1"

#rbinding all 4 groups together + making them a factor
LF_MB_All<-rbind(LF0_MB_P0,LF0_MB_P1,LF1_MB_P0,LF1_MB_P1)

LF_MB_All$LF_PCLO<-as.factor(LF_MB_All$LF_PCLO)

#Pairwise Wilcoxen Rank Sum test on Mutation Burden+LF+PCLO
pairwise.wilcox.test(x=LF_MB_All$Mutation_Burden,g=LF_MB_All$LF_PCLO,data=LF_MB_All,p.adjust.method="bonferroni",paired=FALSE,alternative="two.sided",conf.level=0.95)

#Secondary Outcome analysis for PCLO mutation status+Age+Liver Fibrosis
LF_Age_P0<-subset(x=LiverFibrosis_completeOnly_rest_FullSet,
                 subset=PCLO=="not mutated",
                 select=c("Age","Liver_Fibrosis"))

LF_Age_P1<-subset(x=LiverFibrosis_completeOnly_rest_FullSet,
                 subset=PCLO=="Mutated",
                 select=c("Age","Liver_Fibrosis"))
LF0_Age_P0<-subset(x=LF_Age_P0,
                  subset=Liver_Fibrosis=="0",
                  select=c("Age"))

LF1_Age_P0<-subset(x=LF_Age_P0,
                  subset=Liver_Fibrosis=="1",
                  select=c("Age"))

LF0_Age_P1<-subset(x=LF_Age_P1,
                  subset=Liver_Fibrosis=="0",
                  select=c("Age"))

LF1_Age_P1<-subset(x=LF_Age_P1,
                  subset=Liver_Fibrosis=="1",
                  select=c("Age"))

#Creating column for groups 
LF0_Age_P0$LF_PCLO<-"PCLO_0/LF_0"
LF1_Age_P0$LF_PCLO<-"PCLO_0/LF_1"
LF0_Age_P1$LF_PCLO<-"PCLO_1/LF_0"
LF1_Age_P1$LF_PCLO<-"PCLO_1/LF_1"

#rbinding all 4 groups together + making them a factor
LF_Age_All<-rbind(LF0_Age_P0,LF0_Age_P1,LF1_Age_P0,LF1_Age_P1)

LF_Age_All$LF_PCLO<-as.factor(LF_Age_All$LF_PCLO)

#Pairwise Wilcoxen Rank Sum test on Mutation Burden+LF+PCLO
pairwise.wilcox.test(x=LF_Age_All$Age,g=LF_Age_All$LF_PCLO,data=LF_Age_All,p.adjust.method="bonferroni",paired=FALSE,alternative="two.sided",conf.level=0.95)

#Initial Distribution of Primary Outcomes
require(ggplot2)

#Histogram 1: Mutation Burden+Not Mutated PCLO
ggplot(PcloMtN, aes(x=Mutation_Burden,fill=PCLO,color=PCLO)) +
  geom_histogram(binwidth=30,linetype="dotted",alpha=0.5,color="mediumorchid3",fill="mediumorchid3")+
  geom_density(size=0.75,alpha=0.5,color="black")+
  scale_linetype_manual(values=c(PCLO="dotted"))+
  labs(title="Histogram of Patient's Mutation Burden without a mutation in PCLO",
       x="Mutation Burden",
       y="Density")+
  theme(legend.position="top",
        panel.border=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor=element_blank(),
        panel.background = element_blank(),
        strip.background.x = element_rect("grey83"),
        axis.line = element_line(color="black"),
        plot.subtitle = element_text(size=15, hjust = 0.5),
        plot.title=element_text(size=15,face="bold", hjust = 0.5),
        axis.title=element_text(size=15,hjust = 0.5))+
  scale_fill_manual(values=c(PCLO="mediumorchid3"),
                    labels=c(PCLO="without mutations"))+
  scale_color_manual(values=c(PCLO="mediumorchid3"),
                     labels=c(PCLO="without mutations"))

#Histogram 2: Mutation Burden+Mutated PCLO
ggplot(PcloMtY, aes(x=Mutation_Burden,fill=PCLO,color=PCLO))+
  geom_histogram(binwidth=30,alpha=0.5,fill="slateblue3",color="slateblue3")+
  geom_density(size=0.75,alpha=0.5,color="black")+
  scale_linetype_manual(values=c(PCLO="dotted"))+
  labs(title="Histogram of Patient's Mutation Burden with mutations in PCLO",
       x="Mutation Burden",
       y="Density")+
  theme(legend.position="top",
        panel.border=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor=element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color="black"),
        plot.subtitle = element_text(size=15, hjust = 0.5),
        plot.title=element_text(size=15,face="bold", hjust = 0.5),
        axis.title=element_text(size=15,hjust = 0.5))+
  scale_fill_manual(values=c(PCLO="slateblue3"),
                    labels=c(PCLO="with mutations"))+
  scale_color_manual(values=c(PCLO="slateblue3"),
                     labels=c(PCLO="with mutations"))

#Intial Distrubution of Secondary outcome + PCLO mutation status 

#Histogram 3: Liver Fibrosis occurrence + Not mutated PCLO
ggplot(PcloLFN, aes(x=Liver_Fibrosis, color=Liver_Fibrosis,fill=Liver_Fibrosis)) +
  geom_bar(linetype="dotted",alpha=0.5)+
  scale_linetype_manual(values=c(PCLO="dotted"))+
  labs(title="Histogram of Liver Fibrosis status in patients without mutation in PCLO",
       x="Liver Fibrosis Occurence",
       y="Density")+
  theme(legend.position="top",
        panel.border=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor=element_blank(),
        panel.background = element_blank(),
        strip.background.x = element_rect("grey83"),
        axis.line = element_line(color="black"),
        plot.subtitle = element_text(size=15, hjust = 0.5),
        plot.title=element_text(size=15,face="bold", hjust = 0.5),
        axis.title=element_text(size=15,hjust = 0.5))+
  scale_fill_manual(values=c('2'="mediumorchid3",'1'="slateblue3"),
                    labels=c('2'="without mutation",'1'="with mutation"))+
  scale_color_manual(values=c('2'="mediumorchid3",'1'="slateblue3"),
                     labels=c('2'="without mutation",'1'="with mutation"))

#Histogram 4: Liver Fibrosis+Mutated PCLO
ggplot(PcloLFY, aes(x=Liver_Fibrosis,color=Liver_Fibrosis,fill=Liver_Fibrosis)) +
  geom_bar(linetype="dotted",alpha=0.5)+
  scale_linetype_manual(values=c(PCLO="dotted"))+
  labs(title="Histogram of Liver Fibrosis status in patients with mutation in PCLO",
       x="Liver Fibrosis Occurence",
       y="Density")+
  theme(legend.position="top",
        panel.border=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor=element_blank(),
        panel.background = element_blank(),
        strip.background.x = element_rect("grey83"),
        axis.line = element_line(color="black"),
        plot.subtitle = element_text(size=15, hjust = 0.5),
        plot.title=element_text(size=15,face="bold", hjust = 0.5),
        axis.title=element_text(size=15,hjust = 0.5))+
  scale_fill_manual(values=c('2'="mediumorchid3",'1'="slateblue3"),
                    labels=c('2'="without mutation",'1'="with mutation"))+
  scale_color_manual(values=c('2'="mediumorchid3",'1'="slateblue3"),
                     labels=c('2'="without mutation",'1'="with mutation"))

# Final Violin Plot for Primary Outcome

#Using rbind() to compile data for violin plot 
PcloMt0$PCLO<-"0"
PcloMt1$PCLO<-"1"
PcloMt<-rbind(PcloMt0,PcloMt1)

#Violin plot for Mutation burden and PCLO
ggplot(PcloMt, aes(x=PCLO,y=Mutation_Burden, fill=PCLO,color=PCLO)) +
  geom_violin(scale="count")+
  labs(x="PCLO",
       y="Mutation Burden")+
  theme(legend.position="bottom",
        panel.border=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor=element_blank(),
        panel.background = element_blank(),
        strip.background.x = element_rect("grey83"),
        axis.line = element_line(color="black"),
        plot.subtitle = element_text(size=15, hjust = 0.5),
        plot.title=element_text(size=15,face="bold", hjust = 0.5),
        axis.title=element_text(size=15,hjust = 0.5))+
  scale_fill_manual(values=c('0'="mediumorchid3",'1'="slateblue3"),
                    labels=c('0'="Not Mutated",'1'="Mutated"))+
  scale_color_manual(values=c('0'="mediumorchid3",'1'="slateblue3"),
                     labels=c('0'="Not Mutated",'1'="Mutated"))+
  scale_linetype_manual(values=c('0'="dotted",'1'="solid"), 
                        labels=c('0'="Not Mutated",'1'="Mutated"))
                        
#Subsetting for violin plot
PcloAge0$PCLO<-"0"
PcloAge1$PCLO<-"1"
PcloAge<-rbind(PcloAge0,PcloAge1)

#Violin Plot for Age vs PCLO 
ggplot(PcloAge, aes(x=PCLO,y=Age, fill=PCLO,color=PCLO)) +
  geom_violin(scale="count")+
  labs(x="PCLO",
       y="Age")+
  theme(legend.position="bottom",
        panel.border=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor=element_blank(),
        panel.background = element_blank(),
        strip.background.x = element_rect("grey83"),
        axis.line = element_line(color="black"),
        plot.subtitle = element_text(size=15, hjust = 0.5),
        plot.title=element_text(size=15,face="bold", hjust = 0.5),
        axis.title=element_text(size=15,hjust = 0.5))+
  scale_fill_manual(values=c('0'="mediumorchid3",'1'="slateblue3"),
                    labels=c('0'="Not Mutated",'1'="Mutated"))+
  scale_color_manual(values=c('0'="mediumorchid3",'1'="slateblue3"),
                     labels=c('0'="Not Mutated",'1'="Mutated"))+
  scale_linetype_manual(values=c('0'="dotted",'1'="solid"), 
                        labels=c('0'="Not Mutated",'1'="Mutated"))
                        
```

### Ariana

```{r}
require(moonBook)
require(webr)
library(dplyr)
library(plotly)
library(ggstatsplot)



DataImp <- BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO

# chi-squared test
x=chisq.test(table(DataImp$Survival_Status,DataImp$PCLO), correct = FALSE)
x
#plot(x[[1]]) 

ggbarstats(DataImp, x=PCLO, y=Survival_Status, package = "ggthemes",
           palette = "Purple_Pink_Gray", title = "Alcohol history vs Mutation")



x=chisq.test(table(DataImp$Alcohol_History,DataImp$PCLO), correct = FALSE)
x
#plot(x)

ggbarstats(DataImp, x=PCLO, y=Alcohol_History, package = "ggthemes",
           palette = "Purple_Pink_Gray", title = "Alcohol history vs Mutation")







Vars <- select(DataImp, Survival_Status, Age, Mutation_Burden, Recurrence, Metastasis, PCLO, Alcohol_History)
WithAlc <- subset(Vars, Alcohol_History %in% "1")
WithoutAlc <- subset(Vars, Alcohol_History %in% "0")

Varstable <- table(Vars$Survival_Status, Vars$Alcohol_History)
chisq.test(Varstable)



###############################################################################################


###Survival Status WithAlc
Varstable <- table(Vars$Survival_Status, Vars$Alcohol_History)
chisq.test(Varstable)

ggbarstats(WithAlc, x=PCLO, y=Survival_Status, package = "ggthemes",
           palette = "Purple_Pink_Gray", title = "Those with alcohol history")



table <- table(WithAlc$PCLO, WithAlc$Survival_Status)
table
chisq.test(table)

###Recurrence WithAlc
Varstable <- table(Vars$Recurrence, Vars$Alcohol_History)
chisq.test(Varstable)

ggbarstats(WithAlc, x=PCLO, y=Recurrence, package = "ggthemes",
           palette = "Purple_Pink_Gray", title = "Those with alcohol history")



table <- table(WithAlc$PCLO, WithAlc$Recurrence)
table
chisq.test(table)


###Metastasis WithAlc
Varstable <- table(Vars$Metastasis, Vars$Alcohol_History)
chisq.test(Varstable)

ggbarstats(WithAlc, x=PCLO, y=Metastasis, package = "ggthemes",
           palette = "Purple_Pink_Gray", title = "Those with alcohol history")



table <- table(WithAlc$PCLO, WithAlc$Metastasis)
table
chisq.test(table)


##And now for continuous variables


WithAlcNoMut <- subset(WithAlc, PCLO %in% "not mutated")
WithAlcYesMut <- subset(WithAlc, PCLO %in% "Mutated")

###Mutation Burden WithAlc
var.test(WithAlc$Mutation_Burden, WithoutAlc$Mutation_Burden)
t.test(WithAlc$Mutation_Burden, WithoutAlc$Mutation_Burden, alternative = "two.sided")


var.test(WithAlcNoMut$Mutation_Burden, WithAlcYesMut$Mutation_Burden)
t.test(WithAlcNoMut$Mutation_Burden, WithAlcYesMut$Mutation_Burden, alternative = "two.sided")


###Age WithAlc
var.test(WithAlc$Age, WithoutAlc$Age)
t.test(WithoutAlc$Age,WithAlc$Age, alternative = "two.sided")
###SIGNIFICANT???
```

```{r}
boxplot(WithAlc$Age, WithoutAlc$Age, ylab = "Age", names = c("Alcohol History","NO Alcohol History"), main = "Alcohol History vs. Age")

Violin_plot <- ggplot(data = DataImp,aes(x = Alcohol_History, y = Age,fill=Age))+
  geom_violin(trim=FALSE)+
  geom_boxplot(width=0.1) #+
  #rotate_x_text(angle = 45)

Violin_plot + theme(
  panel.background = element_rect(fill = "mediumorchid3",
                                  colour = "mediumorchid3",
                                  size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                  colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                  colour = "white")
)



var.test(WithAlcNoMut$Age, WithAlcYesMut$Age)
t.test(WithAlcNoMut$Age, WithAlcYesMut$Age, alternative = "two.sided")



#####NOW WITHOUTALC #######

###Survival Status WithoutAlc
ggbarstats(WithoutAlc, x=PCLO, y=Survival_Status, package = "ggthemes",
           palette = "Purple_Pink_Gray", title = "Those WITHOUT alcohol history")



table <- table(WithoutAlc$PCLO, WithoutAlc$Survival_Status)
table
chisq.test(table)

###Recurrence WithoutAlc
ggbarstats(WithoutAlc, x=PCLO, y=Recurrence, package = "ggthemes",
           palette = "Purple_Pink_Gray", title = "Those WITHOUT alcohol history")



table <- table(WithoutAlc$PCLO, WithoutAlc$Recurrence)
table
chisq.test(table)


###Metastasis WithoutAlc
ggbarstats(WithoutAlc, x=PCLO, y=Metastasis,package = "ggthemes",
           palette = "Purple_Pink_Gray", title = "Those WITHOUT alcohol history")



table <- table(WithoutAlc$PCLO, WithoutAlc$Metastasis)
table
chisq.test(table)


##And now for continuous variables


WithoutAlcNoMut <- subset(WithoutAlc, PCLO %in% "not mutated")
WithoutAlcYesMut <- subset(WithoutAlc, PCLO %in% "Mutated")

###Mutation Burden WithAlc
var.test(WithoutAlcNoMut$Mutation_Burden, WithoutAlcYesMut$Mutation_Burden)
t.test(WithoutAlcNoMut$Mutation_Burden, WithoutAlcYesMut$Mutation_Burden, alternative = "two.sided")


###Age WithAlc
var.test(WithoutAlcNoMut$Age, WithoutAlcYesMut$Age)
t.test(WithoutAlcNoMut$Age, WithoutAlcYesMut$Age, alternative = "two.sided")
```

### Tri

```{r}
require(readr)
require(dplyr)
require(ggplot2)
require(ggpubr)
require(pastecs)
require(ggstatsplot)
library(graphics)
library(corrplot)
require(tableone)
require(sjPlot)


### Using the imputed data
#load("Main_variables_imputed_dataset_MidtermProject_02_13_22.RData")

Midtermdata <- BurdenMut_HepatitisHis_HisAlc_BMI_imputed_New_Full_PCLO

Midtermdata$Metastasis <- factor(Midtermdata$Metastasis)

Metattab <- table(Midtermdata$Metastasis,Midtermdata$PCLO)

chisq.test(Metattab,correct = F)

Mettabplot <- ggplot(Midtermdata, aes(x=Metastasis,fill=PCLO))+
  geom_bar(position="dodge", color="black")+
  scale_fill_manual(values = c("mediumorchid3","slateblue3"))+
  labs(title ="Metastasis history by PCLO mutation",
       y="Frequency") +
  theme (plot.title = element_text(hjust = 0.5, size= 15, face="bold"),
         panel.background = element_blank(),
         strip.text.x = element_text(colour = "black", size = 15),
         axis.line = element_line(colour = "black"),
         axis.title.x = element_text(colour = "black", size = 10,face="bold"),
         axis.text.x = element_text(colour = "black", size = 8, face = "bold"),
         axis.title.y = element_text(colour = "black", size = 10,face="bold"),
         axis.text.y = element_text(colour = "black", size = 10, face = "bold"))

# length of palette giving issues, using default colors
MetatGraph <- ggbarstats(Midtermdata,x = Metastasis, y=PCLO)

#View(paletteer::palettes_d_names)

#SEX

Sextab <- table(Midtermdata$Sex,Midtermdata$PCLO)

Chi <- chisq.test(Sextab)
Chi$statistic
fisher.test(Sextab)

Sextabplot <- ggplot(Midtermdata, aes(x=Sex,fill=PCLO))+
  geom_bar(position="dodge", color="black")+
  scale_fill_manual(values = c("mediumorchid3","slateblue3"))+
  labs(title ="Sex by PCLO mutation",
       y="Frequency") +
  theme (plot.title = element_text(hjust = 0.5, size= 15, face="bold"),
         panel.background = element_blank(),
         strip.text.x = element_text(colour = "black", size = 15),
         axis.line = element_line(colour = "black"),
         axis.title.x = element_text(colour = "black", size = 10,face="bold"),
         axis.text.x = element_text(colour = "black", size = 8, face = "bold"),
         axis.title.y = element_text(colour = "black", size = 10,face="bold"),
         axis.text.y = element_text(colour = "black", size = 10, face = "bold"))


SexGraph <- ggbarstats(Midtermdata,x = Sex, y=PCLO)

### Compare Metastasis w/ Sex

Metatandsextab <- table(Midtermdata$Metastasis,Midtermdata$Sex)
chisq.test(Metatandsextab,correct = F)

#Female

FemaleData <- filter(Midtermdata,Sex=="FEMALE")
MaleData <- filter(Midtermdata, Sex=="MALE")

MetastasisandFe <- table(FemaleData$Metastasis,FemaleData$PCLO)
fisher.test(MetastasisandFe) #Not significant

#Male

MetastasisandMale <- table(MaleData$Metastasis,MaleData$PCLO)
chisq.test(MetastasisandMale,correct = F) #Not significant


### Compare Recurrence w/ Sex

Recurandsextab <- table(Midtermdata$Sex,Midtermdata$Recurrence)
chisq.test(Recurandsextab,correct = F)

#Female

RecurFeTab <- table(FemaleData$PCLO,FemaleData$Recurrence)
fisher.test(RecurFeTab) #Not significant

#Male

RecurMaleTab <- table(MaleData$PCLO,MaleData$Recurrence)
chisq.test(RecurMaleTab,correct = F) #Not significant

### Compare Survival Status w/ Sex

Surviveandsextab <- table(Midtermdata$Sex,Midtermdata$Survival_Status)
chisq.test(Surviveandsextab,correct = F) #SIGNIFICANT

#Female

SurviveFeTab <- table(FemaleData$PCLO,FemaleData$Survival_Status)
fisher.test(SurviveFeTab) #Not significant


#Male

SurviveMaleTab <- table(MaleData$PCLO,MaleData$Survival_Status)
chisq.test(SurviveMaleTab,correct = F)

### Compare Age w/ Sex

var.test(FemaleData$Age,MaleData$Age)
t.test(FemaleData$Age,MaleData$Age)


#Female

FemaleMutation <- filter(FemaleData,PCLO=="Mutated")
FemaleNotMutated <- filter(FemaleData, PCLO=="not mutated")

t.test(FemaleMutation$Age,FemaleNotMutated$Age) #Not significant

#Male

MaleMutation <- filter(MaleData,PCLO=="Mutated")
MaleNotMutated <- filter(MaleData, PCLO=="not mutated")

t.test(MaleMutation$Age,MaleNotMutated$Age) #Not significant

### Compare Mutation Burden w/ Sex

t.test(FemaleData$Mutation_Burden,MaleData$Mutation_Burden) #SIGNIFICANCE

#Female

t.test(FemaleMutation$Mutation_Burden,FemaleNotMutated$Mutation_Burden) #Not significant

#Male

t.test(MaleMutation$Mutation_Burden,MaleNotMutated$Mutation_Burden) #Not significant
```


### Paul

```{r}
#python setup
#use conda_env.yaml to generate conda env names 'BSR1026_midterm' before running
Sys.setenv(RETICULATE_PYTHON = "/Users/appankow/opt/anaconda3/envs/BSR1026_midterm/bin/python")
library(reticulate)
use_condaenv("BSR1026_midterm", required = T)
```


```{python}
import statistics
import numpy as np
import pandas as pd
import seaborn as sns
import matplotlib.pyplot as plt
from bioinfokit.analys import stat

df = pd.read_csv('data/Midterm_Dataset.csv')
print(df)

sns.set_theme(style="ticks", color_codes=True)
custom_palette = ['#B452CD','#6959CD']
sns.set_palette(custom_palette)

sns.catplot(x="BMI_Category", y="Age", hue="PCLO", kind="swarm", data=df)
sns.catplot(x="Recurrence", y="Age", hue="PCLO", kind="swarm", data=df)
sns.catplot(x="PCLO", y="Age", hue="Recurrence", kind="swarm", data=df)

#Chi squared test
ctab = pd.crosstab(df.PCLO,df.Recurrence)
print(ctab)

res = stat()
res.chisq(df=ctab)
print(res.summary)

print(res.expected_df)

#Chi squared test
ctab2 = pd.crosstab(df.PCLO,df.BMI_Category)
print(ctab2)

res2 = stat()
res2.chisq(df=ctab)
print(res.summary)

print(res2.expected_df)

# cross-validate with Fisher
from scipy.stats import fisher_exact

#Fisher test for Crosstab PCLO and Recurrence
fisher_exact(ctab)

#Fisher test for Crosstab PCLO and BMI
pvalue = fisher_exact(ctab)
print(pvalue)
```

```{python}
sns.set_theme(style="ticks", color_codes=True)
custom_palette = ['#B452CD','#6959CD']
sns.set_palette(custom_palette)

sns.catplot(x="BMI_Category", y="Age", hue="PCLO", kind="swarm", data=df)
sns.catplot(x="Recurrence", y="Age", hue="PCLO", kind="swarm", data=df)
sns.catplot(x="PCLO", y="Age", hue="Recurrence", kind="swarm", data=df)

#Chi squared test
ctab = pd.crosstab(df.PCLO,df.Recurrence)
print(ctab)

res = stat()
res.chisq(df=ctab)
print(res.summary)

print(res.expected_df)

#Chi squared test
ctab2 = pd.crosstab(df.PCLO,df.BMI_Category)
print(ctab2)

res2 = stat()
res2.chisq(df=ctab)
print(res.summary)

print(res2.expected_df)

# cross-validate with Fisher
from scipy.stats import fisher_exact

#Fisher test for Crosstab PCLO and Recurrence
fisher_exact(ctab)

#Fisher test for Crosstab PCLO and BMI
pvalue = fisher_exact(ctab)
print(pvalue)
```

### Alec

```{r, message = F, warning = F}
# imports
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggstatsplot)


# read in data
tcga <- readxl::read_xlsx("data/Midterm_Dataset.xlsx") %>%
  mutate(
    PCLO = if_else(PCLO == "Mutated", "mutated", "unmutated", missing = NULL),
    PCLO_Status = if_else(PCLO == "mutated", 1, 0, missing = NULL)
  )
head(tcga)
```

### Data cleaning / imputation

For variables where missing data makes up greater than 33 observations, omit NA observations from downstream analysis.
For variables with fewer missing observations, replace NAs with the most frequent category or the sample mean. 

```{r}
# with 'pseudo-imputation' for PCLO status / Hepatitis history
tcga.imp <- tcga %>% 
  mutate(
    PCLO = ifelse(is.na(PCLO), "unmutated", PCLO),
    Hepatitis_History = ifelse(is.na(Hepatitis_History), 0, Hepatitis_History)
    )

tcga.imp
```



```{r}
tcga.imp %>% group_by(Race) %>% summarise(n = n())
```


## Primary variable: Recurrence

```{r}
PCLO.recurrence.sum <- tcga.imp %>% 
  select(PCLO, Recurrence) %>%
  group_by(PCLO, Recurrence) %>%
  na.omit() %>% 
  summarise(n = n()) 
PCLO.recurrence.sum
```

```{r}
mat <- matrix(PCLO.recurrence.sum$n, nrow = 2)
fisher.test(mat)
```

```{r}
# unmutated: mediumorchid3; mutated: slateblue3 
cols <- c("slateblue3", "mediumorchid3")

ggplot(tcga, aes(factor(Recurrence), fill = PCLO)) + 
  geom_bar(position = "fill") +
  theme_classic() + 
  geom_label(stat = "count", aes(label=..count..), position = position_fill(vjust = 0.5)) +
  scale_fill_manual(values = cols) +
  labs(x = "Recurrence", y = "Proportion") +
  coord_flip() + ggplot2::theme(legend.position = "bottom") 
#+ facet_wrap(~ Hepatitis_History, labeller = label_both)

ggsave("figs/recurrence_desc_bars.png", height = 3, width = 6, dpi = 1000)
```

## Secondary variable: Hepatitis_History

#### Stratified analysis

Primary research variables

- age (continuous)
- mutation burden (continuous)

- metastisis
- recurrence
- survival status

My secondary research variable

- hepatitis history

```{r}
# filtered dfs for tests... 
tcga.mut <- tcga.imp %>% filter(PCLO_Status == 1)
tcga.norm <- tcga.imp %>% filter(PCLO_Status == 0)

tcga.mut.hep <- tcga.mut %>% filter(Hepatitis_History == 1)
tcga.norm.hep <- tcga.norm %>% filter(Hepatitis_History == 1)

tcga.mut.nohep <- tcga.mut %>% filter(Hepatitis_History == 0)
tcga.norm.nohep <- tcga.norm %>% filter(Hepatitis_History == 0)
```

```{r}
ggplot(tcga, aes(factor(Hepatitis_History), fill = PCLO)) + 
  geom_bar(position = "fill") +
  theme_classic() + 
  coord_flip() +
  geom_label(stat = "count", aes(label=..count..), position = position_fill(vjust = 0.5)) +
  scale_fill_manual(values = cols) +
  labs(x = "Hepatitis History", y = "Proportion")

ggsave("figs/hep_hist_all_barplot.png", height = 3, width = 8, dpi = 1000)
```

### Age

```{r}
ggplot(tcga.imp, aes(Age, fill = PCLO)) + 
  geom_histogram(binwidth = 5) +
  theme_classic() + 
  scale_fill_manual(values = cols) + 
  facet_grid(PCLO ~ Hepatitis_History, labeller = label_both) + theme(legend.position = "none") +
  labs(y = "Count")

ggsave("figs/Age_hep_hist_strat.png", height = 4, width = 5, dpi = 1000)
```

```{r}
var = "Age"
# unstratified by hepatitis hist
t1 <- t.test(tcga.mut[var], tcga.norm[var])
t1
```


```{r}
var.test(tcga.mut$Age, tcga.norm$Age)
```

```{r}
# with hepatitis hist
t.test(tcga.mut.hep[var], tcga.norm.hep[var], var.equal = F)
```

Also finds an association with non-parametric test...

```{r}
wilcox.test(tcga.mut.hep$Age, tcga.norm.hep$Age, alternative = "two.sided")
```

```{r}
# with no hepatitis hist
t.test(tcga.mut.nohep[var], tcga.norm.nohep[var])
```

```{r}
# with no hepatitis hist
var.test(tcga.mut.nohep$Age, tcga.norm.nohep$Age)
```

### Mutation burden

```{r}
ggplot(tcga.imp, aes(log10(Mutation_Burden + 1), fill = PCLO)) + 
  geom_histogram() +
  theme_classic() + 
  scale_fill_manual(values = cols) + 
  facet_grid(PCLO ~ Hepatitis_History, labeller = label_both) + theme(legend.position = "none") +
  labs(y = "Count")

ggsave("figs/mutation_hep_hist_strat.png", height = 3, width = 8, dpi = 1000)
```

```{r}
var = "Mutation_Burden"
# unstratified by hepatitis hist
t.test(tcga.mut[var], tcga.norm[var])
```

```{r}
# with hepatitis hist
t.test(tcga.mut.hep[var], tcga.norm.hep[var])
```

```{r}
# with no hepatitis hist
t.test(tcga.mut.nohep[var], tcga.norm.nohep[var])
```

### Recurrence

```{r}
ggplot(tcga.imp, aes(factor(Recurrence), fill = PCLO)) + 
  geom_bar(position = "fill") +
  theme_classic() + 
  coord_flip() +
  geom_label(stat = "count", aes(label=..count..), position = position_fill(vjust = 0.5)) +
  scale_fill_manual(values = cols) +
  labs(x = "Recurrence", y = "Proportion") + facet_wrap(~ Hepatitis_History, labeller = label_both)

ggsave("figs/recurrence_hep_hist_strat_bars.png", height = 3, width = 8, dpi = 1000)
```

```{r}
# unstratified by hepatitis hist
tcga.sum <- tcga.imp %>% group_by(PCLO, Recurrence) %>%
  summarise(n = n()) %>% na.omit()
tcga.sum

m <- matrix(tcga.sum$n, nrow = 2)
fisher.test(m)
```

```{r}
# with hepatitis hist.
tcga.sum <- tcga.imp %>% 
  filter(Hepatitis_History == 1) %>%
  group_by(PCLO, Recurrence) %>%
  summarise(n = n()) %>% na.omit()
tcga.sum

m <- matrix(tcga.sum$n, nrow = 2)
fisher.test(m)
```

```{r}
# without hepatitis hist.
tcga.sum <- tcga.imp %>% 
  filter(Hepatitis_History == 0) %>%
  group_by(PCLO, Recurrence) %>%
  summarise(n = n()) %>% na.omit()
tcga.sum

m <- matrix(tcga.sum$n, nrow = 2)
fisher.test(m)
```

### Metastisis

```{r}
ggplot(tcga.imp, aes(factor(Metastasis), fill = PCLO)) + 
  geom_bar(position = "fill") +
  theme_classic() + 
  coord_flip() +
  geom_label(stat = "count", aes(label=..count..), position = position_fill(vjust = 0.5)) +
  scale_fill_manual(values = cols) +
  labs(x = "Metastasis", y = "Proportion") + facet_wrap(~ Hepatitis_History, labeller = label_both)

ggsave("figs/metastasis_hep_hist_strat_bars.png", height = 3, width = 8, dpi = 1000)
```

```{r}
# unstratified by hepatitis hist
tcga.sum <- tcga %>% group_by(PCLO, Metastasis) %>%
  summarise(n = n()) %>% na.omit()
tcga.sum

m <- matrix(tcga.sum$n, nrow = 2)
fisher.test(m)
```

```{r}
# with hepatitis hist.
tcga.sum <- tcga %>% 
  filter(Hepatitis_History == 1) %>%
  group_by(PCLO, Metastasis) %>%
  summarise(n = n()) %>% na.omit()
tcga.sum

m <- matrix(tcga.sum$n, nrow = 2)
fisher.test(m)
```


```{r}
# without hepatitis hist.
tcga.sum <- tcga %>% 
  filter(Hepatitis_History == 0) %>%
  group_by(PCLO, Metastasis) %>%
  summarise(n = n()) %>% na.omit()
tcga.sum

m <- matrix(tcga.sum$n, nrow = 2)
fisher.test(m)
```

### Survival status

```{r}
ggplot(tcga.imp, aes(factor(Survival_Status), fill = PCLO)) + 
  geom_bar(position = "fill") +
  theme_classic() + 
  coord_flip() +
  geom_label(stat = "count", aes(label=..count..), position = position_fill(vjust = 0.5)) +
  scale_fill_manual(values = cols) +
  labs(x = "Survival_Status", y = "Proportion") + facet_wrap(~ Hepatitis_History, labeller = label_both)

ggsave("figs/survival_hep_hist_strat_bars.png", height = 3, width = 8, dpi = 1000)
```

```{r}
# unstratified by hepatitis hist
tcga.sum <- tcga %>% group_by(PCLO_Status, Survival_Status) %>%
  summarise(n = n()) %>% na.omit()
tcga.sum

m <- matrix(tcga.sum$n, nrow = 2)
fisher.test(m)
```

```{r}
# with hepatitis hist.
tcga.sum <- tcga %>% 
  filter(Hepatitis_History == 1) %>%
  group_by(PCLO, Survival_Status) %>%
  summarise(n = n()) %>% na.omit()
tcga.sum

m <- matrix(tcga.sum$n, nrow = 2)
fisher.test(m)
```

```{r}
# without hepatitis hist.
tcga.sum <- tcga %>% 
  filter(Hepatitis_History == 0) %>%
  group_by(PCLO_Status, Survival_Status) %>%
  summarise(n = n()) %>% na.omit()
tcga.sum

m <- matrix(tcga.sum$n, nrow = 2)
fisher.test(m)
```
